Learning action models with minimal observability

### 2019-05-01 doc/aij18/minor-revision/learning_action_models.tex
AIJ18 minor revision folder created

\textcolor[rgb]{1.00,0.00,0.00}{Our proposal to address a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to transform $\Lambda$ into a planning task $P_{\Lambda}$. The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that build the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the actions and states of the observed plan trace $\tau = \tup {s_0, \ldots, s_n}$. Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition, a positive effect or a negative effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\tau$, including that the final reached state comprises $s_n$. Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states of $\tau$.}

@@

We are now ready to define the propositional encoding of the \textcolor[rgb]{1.00,0.00,0.00}{model fluents} of $pre(\xi)$, $del(\xi)$ and $add(\xi)$. For every $\xi$ and $p\in \Psi_{\xi}$, we create:

\begin{itemize}
\item $pre_p(\xi)$: \textcolor[rgb]{1.00,0.00,0.00}{model fluent} formed by the combination of the prefixes {\small \texttt{pre}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g. {\small \texttt{pre\_stack\_on\_v1\_v2}}, for $name(\xi)={\small\texttt{stack}}$ and $p={\small\texttt{on($v_1,v_2$)}}$)
\item $del_p(\xi)$: \textcolor[rgb]{1.00,0.00,0.00}{model fluent} formed by the combination of the prefixes {\small \texttt{del}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g. {\small  \texttt{del\_stack\_on\_v1\_v2}})
\item $add_p(\xi)$: \textcolor[rgb]{1.00,0.00,0.00}{model fluent} formed by the combination of the prefixes {\small \texttt{add}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g. {\small \texttt{add\_stack\_on\_v1\_v2}})
\end{itemize}

For a given action model $\xi$, if a fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ holds in a state, it means that $p$ is a precondition/negative/positive effect of $\xi$. For instance, Figure~\ref{fig:encodedstack} shows the conjunction of \textcolor[rgb]{1.00,0.00,0.00}{model fluents} that represents the propositional encoding of the preconditions, negative effects and positive effects of the action model corresponding to the {\em stack} operator shown in Figure~\ref{fig:stack}.

@@

%\subsection{Learning from observations of plan executions}
Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$, with $\tau$ formed by an $n$-action sequence $\tup{a_1, \ldots, a_n}$ and a $m$-state trajectory $\tup{s_0, s_1, \ldots, s_m}$ ($\tau = \langle s_0, a_1, \ldots, a_n, s_m \rangle$), the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:

@@

\item $F_{\Lambda}$ \textcolor[rgb]{1.00,0.00,0.00}{extends $F$ with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of $\tau$}. Specifically, $F_{\Lambda}$ contains:
\begin{itemize}
\item The set of fluents obtained from $s_0$; i.e., $F$.
\item The \textcolor[rgb]{1.00,0.00,0.00}{model fluents} $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section \ref{propositional_encoding}
\item A set of fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to represent the $i^{th}$ observable action of $\tau$. In the example of Figure~\ref{fig:example-plans}, the two observed actions {\small \texttt{(putdown B)}} and {\small \texttt{(stack  A  B)}} would be encoded as fluents  {\small \texttt{(plan-putdown B i1)}} and {\small \texttt{(plan-stack A B i2)}} to indicate that {\small \texttt{(putdown B)}} is observed in the first place and {\small \texttt{(stack  A  B)}} is the second observed action.
\item Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions of $\tau$. The former is used to ensure that actions are executed in the same order as they are observed in $\tau$. The latter is used to iterate to the next planning step when solving $P_{\Lambda}$.
\item A set of fluents $\{test_j\}_{0\leq j\leq m}$, to point at the state observation $s_j\in\tau$ where the action model is
validated. In the example of Figure~\ref{fig:example-plans} two tests are required to validate the programmed action model, one test at $s_0$ and another one at $s_4$.
\item A fluent, $mode_{prog}$, to indicate whether action models are being programmed or validated.
\end{itemize}

@@

\item $A_{\Lambda}$ includes \textcolor[rgb]{1.00,0.00,0.00}{three types of actions that give rise to the actions of $\pi_\Lambda$}.
\begin{enumerate}
\item Actions for \textcolor[rgb]{1.00,0.00,0.00}{{\em inserting} a component (precondition, positive effect or negative effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models}. These actions will form the prefix of the solution plan $\pi_\Lambda$. Among the \emph{inserting} actions, we find:
\begin{itemize}
\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$. \textcolor[rgb]{1.00,0.00,0.00}{A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $del_p$ nor $add_p$ exist in $\xi$.}

@@

\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$. \textcolor[rgb]{1.00,0.00,0.00}{A positive effect is inserted in $\xi$ under the same conditions of a precondition insertion, and a negative effect is inserted in $\xi$ when neither $del_p$ nor $add_p$ appear in $\xi$ but $pre_p$ does.}

@@

\textcolor[rgb]{1.00,0.00,0.00}{For instance, given $name(\xi)=${\tt{\small stack}} and $C_{pre-stack}=\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_holding\_v2), (pre\_stack\_on\_v1\_v2),(pre\_stack\_clear\_v1),(pre\_stack\_clear\_v1),}}$\ldots \}$, the insertion of each item $c \in C_{pre-stack}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$ as long as $c \notin pre(\xi)$,  $c \notin add(\xi)$ and $c \notin del(\xi)$. The same applies to effects with respect to sets $C_{add-stack}$ and $C_{del-stack}$ that would include all fluents starting with prefix  {\tt{\small add}} and {\tt{\small del}}, respectively.}

@@

\textcolor[rgb]{1.00,0.00,0.00}{Note that executing an insert action, e.g.{\tt{\small (insert\_pre\_stack\_holding\_v1)}}, will add the corresponding model fluent {\tt{\small (pre\_stack\_holding\_v1)}} to the successor state. Hence, the execution of the insert actions of $\pi_\Lambda$ yield a state containing the valuation of the model fluents that shape every $\xi \in \mathcal{M}$. For example, executing the insert actions that shape the action model $name(\xi)=${\tt{\small putdown}} leads to a state containing the positive literals {\tt{\small (pre\_putdown\_holding\_v1),(eff\_putdown\_holding\_v1),\\ (eff\_putdown\_clear\_v1),
(eff\_putdown\_ontable\_v1),(eff\_putdown\_handempty)}}.}

@@

    These actions will be part of the postfix of the plan $\pi_\Lambda$ and \textcolor[rgb]{1.00,0.00,0.00}{they determine the application of the learned action models according to the values of the model fluents in the current state configuration}. Figure~\ref{fig:compilation} shows the PDDL encoding of {\tt{\small (apply\_stack)}} for applying the action model of the {\em stack} operator. \textcolor[rgb]{1.00,0.00,0.00}{Let's assume the action {\tt{\small (apply\_stack blockB blockA)}} is in $\pi_\Lambda$. Executing this action in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$. For example, if  $\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}}$\} \subset s$ then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$. Otherwise, a different set of precondition literals will be checked. The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.}

\textcolor[rgb]{1.00,0.00,0.00}{    Note that executing an apply action, e.g.{\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if $name(\xi)=${\tt{\small stack}} has been correctly programmed by the insert actions. Hence, while \textbf{insert actions} add the values of the \textbf{model fluents} that shape $\xi$, the \textbf{apply actions} add the values of the \textbf{fluents of $F$} that result from the execution of $\xi$.}

@@

\textcolor[rgb]{1.00,0.00,0.00}{There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$. The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$.}

@@

%As shown in Figure \ref{fig:plan-lplan}, the learned action models $\mathcal{M'}$ can be easily inferred from the solution plan $\pi_\Lambda$, just by looking at the programming actions. However, for the shake of formalizing the extraction of the learned models, we will infer them from the final state of the solution plan $s_g=\theta(I_\Lambda,\pi_\Lambda)$ by looking at the values for the fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$ which encode the preconditions and effects of the validated action models. For each $\xi \in \mathcal{M'}$, we build the sets of preconditions, positive effects and negative effects as follows:

@@

The logical inference process our approach is based on has trouble learning preconditions that do not appear as negative effects since in this case no change is observed between the pre-state and post-state of an action. This is specially relevant for static predicates that never change and, hence, only appear as preconditions in the actions. In order to address this shortcoming \textcolor[rgb]{1.00,0.00,0.00}{and complete the list of learned preconditions}, we apply a post-process based on the one proposed in~\cite{kuvcera2018louga}. \textcolor[rgb]{1.00,0.00,0.00}{The idea lies in going through every action and counting the number of cases where a literal is present before the action is executed and the number of cases where it is not present. If a literal is present in all the cases before the action, the literal is considered to be a precondition.}

\textcolor[rgb]{1.00,0.00,0.00}{In order to obtain a complete trace, the proposal in ~\cite{kuvcera2018louga} applies the sequence of actions of the input trace and infers the preconditions from this \FO action sequence}. In our case, since the sequence of actions of the input trace might not be fully observable, we produce the traces by applying the actions found in the validation part of the solution plan. For instance, in the example of the figure \ref{fig:plan-lplan}, the sequence of actions used to produce the complete trace would be {\tt{\small(unstack blockB blockA)}}, {\tt{\small(put-down blockB)}}, {\tt{\small(pick-up blockA)}}, and {\tt{\small(stack blockA blockB)}}.

%This post-process allows \FAMA to learn more complete lists of preconditions and deal with the always problematic static predicates.

@@

%Given a plan $\pi_\Lambda$ that solves $P_\Lambda$, and the learned action models $\mathcal{M'}$ that solve $\Lambda=\tup{\mathcal{M}, \tau}$, \FAMA contributes with the following compilation properties:

@@

\begin{proof}[Proof sketch]
\begin{small}
  By definition, $\Psi_{\xi}\subseteq \Psi_v$ fully captures the set of elements that can appear in an action model $\xi\in\mathcal{M}$. The compilation does not discard any possible set of action models $\mathcal{M}'$ definable within $\Psi_v$ that satisfies the observed state trajectory and action sequence of $\tau$. This means that for every $\mathcal{M}'$ that solves $\Lambda$, there exists a plan $\pi_\Lambda$ that can be built selecting the appropriate programming, apply and validate actions from the $P_{\Lambda}$ compilation.
  %a solution plan $\pi_\Lambda$ can be built selecting the corresponding  $\mathsf{programPre_{p,\xi}}$ and $\mathsf{programEff_{p,\xi}}$ actions according to $\mathcal{M}'$ and later, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ and $\mathsf{validate_{i}}$ actions according to $\tau$.
\end{small}
\end{proof}

@@

%\subsection{Optimizing the compilation with background knowledge}
%A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn logic programs from data~\cite{muggleton1994inductive}. Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case to optimize the performance of the $P_{\Lambda}$ compilation.

%\subsubsection{Static predicates}
%A {\em static predicate} $p \in \Psi_\Lambda$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}. Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete. Given a static predicate $p$:
%\begin{itemize}
%\item Fluents $del_p(\xi)$ and $add_p(\xi)$, such that $p\in \Psi_v$ is the static predicate, can be discarded for every $\xi\in\mathcal{M}$.
%\item Actions $\mathsf{programEff_{p,\xi}}$ (s.t. $p\in \Psi_v$ is a parameterized static predicate) can also be discarded for every $\xi\in\mathcal(M)$.
%\end{itemize}

%Static predicates can also constrain the space of possible preconditions by looking at the given set of state observations in $\tau$. One can assume that if a precondition $p\in \Psi_v$ (s.t. $p\in \Psi_v$ is a static predicate parameterized with $\Omega_v$) is not compliant with the observations in $\tau$ then, fluents $pre_p(\xi)$ and actions $\mathsf{programPre_{p,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$. For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\tau$.

%Furthermore looking as well at the given plan traces, fluents $pre_p(\xi)$ and actions $\mathsf{programPre_{p,\xi}}$ are also discardable for every $\xi\in\mathcal(M)$ if a precondition $p\in \Psi_v$ (s.t. $p\in \Psi_v$ is a static predicate parameterized with $\Omega_v$) is not possible according to $\tau$. Back to the {\em zenotravel} domain, if a plan trace $\tau$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.

%\subsubsection{State constraints}
%The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.  If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states. For instance, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.

%{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}. Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.

%The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$. For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}. There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.

%\begin{figure}[hbt!]
%  \begin{footnotesize}
%    \begin{center}
%$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
%$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
%$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
%$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
%$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
%    \end{center}
%\end{footnotesize}
% \caption{\small An example of the strongest invariant for the {\em blocksworld} domain.}
%\label{fig:strongest-invariant}
%\end{figure}

%A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}. For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a mutex because $block_A$ can only be on top of a single block.

%A {\em domain invariant} is an instance-independent invariant, i.e. holds for any possible initial state and set of objects. Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state. Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}. For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.

%An interesting contribution of our compilation is that the validation of an action model can also be done with {\em state constraints}. Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check also that the learned model satisfy a constraint $\phi\in\Phi$:
%\begin{small}
%\begin{align*}
%\hspace*{7pt}\pre(\mathsf{validate_{i}})=&\phi\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\Phi|}\cup \{mode_{val}\},\\
%\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
%\end{align*}
%\end{small}

%This redefinition of validate actions applies also to {\em trajectory constraints} because LTL formulae can be represented using classical action preconditions and goals by encoding the Non-deterministic B\"{u}chi Automaton (NBA), that is equivalent to the corresponding LTL formula, as part of the classical planning tasks~\cite{baier2006planning}.

%Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan. Constraints can be used to confine earlier the set of possible \strips\ action models and reduce then the learning hypothesis space. With regard to our compilation, {\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema. For example given the {\em domain mutex} $\phi=(\neg f_1\vee \neg f_2)$ such that $f_1\in F_v(\xi)$ and $f_2\in F_v(\xi)$, we can redefine the corresponding programming actions for {\bf removing} the {\em precondition} $f_1\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$ as:

%\begin{small}
%\begin{align*}
%\hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
%\cond(\mathsf{programPre_{f_1,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
%\end{align*}
%\end{small}

>>>>

Our proposal to address a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to transform $\Lambda$ into a planning task $P_{\Lambda}$. The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that build the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the actions and states of the observed plan trace $\tau = \tup {s_0, \ldots, s_n}$. Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition, a positive effect or a negative effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\tau$, including that the final reached state comprises $s_n$. Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states of $\tau$.

@@

We are now ready to define the propositional encoding of the model fluents of $pre(\xi)$, $del(\xi)$ and $add(\xi)$. For every $\xi$ and $p\in \Psi_{\xi}$, we create:

\begin{itemize}
\item $pre_p(\xi)$: model fluent formed by the combination of the prefixes {\small \texttt{pre}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g. {\small \texttt{pre\_stack\_on\_v1\_v2}}, for $name(\xi)={\small\texttt{stack}}$ and $p={\small\texttt{on($v_1,v_2$)}}$)
\item $del_p(\xi)$: model fluent formed by the combination of the prefixes {\small \texttt{del}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g. {\small  \texttt{del\_stack\_on\_v1\_v2}})
\item $add_p(\xi)$: model fluent formed by the combination of the prefixes {\small \texttt{add}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g. {\small \texttt{add\_stack\_on\_v1\_v2}})
\end{itemize}

For a given action model $\xi$, if a fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ holds in a state, it means that $p$ is a precondition/negative/positive effect of $\xi$. For instance, Figure~\ref{fig:encodedstack} shows the conjunction of model fluents that represents the propositional encoding of the preconditions, negative effects and positive effects of the action model corresponding to the {\em stack} operator shown in Figure~\ref{fig:stack}.

@@

Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$, with $\tau$ formed by an $n$-action sequence $\tup{a_1, \ldots, a_n}$ and a $m$-state trajectory $\tup{s_0, s_1, \ldots, s_m}$ ($\tau = \langle s_0, a_1, \ldots, a_n, s_m \rangle$), the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:

@@

\item $F_{\Lambda}$ extends $F$ with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of $\tau$. Specifically, $F_{\Lambda}$ contains:
\begin{itemize}
\item The set of fluents obtained from $s_0$; i.e., $F$.
\item The model fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section \ref{propositional_encoding}
\item A set of fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to represent the $i^{th}$ observable action of $\tau$. In the example of Figure~\ref{fig:example-plans}, the two observed actions {\small \texttt{(putdown B)}} and {\small \texttt{(stack  A  B)}} would be encoded as fluents  {\small \texttt{(plan-putdown B i1)}} and {\small \texttt{(plan-stack A B i2)}} to indicate that {\small \texttt{(putdown B)}} is observed in the first place and {\small \texttt{(stack  A  B)}} is the second observed action.
\item Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions of $\tau$. The former is used to ensure that actions are executed in the same order as they are observed in $\tau$. The latter is used to iterate to the next planning step when solving $P_{\Lambda}$.
\item A set of fluents $\{test_j\}_{0\leq j\leq m}$, to point at the state observation $s_j\in\tau$ where the action model is
validated. In the example of Figure~\ref{fig:example-plans} two tests are required to validate the programmed action model, one test at $s_0$ and another one at $s_4$.
\item A fluent, $mode_{prog}$, to indicate whether action models are being programmed or validated.
\end{itemize}

@@

\item $A_{\Lambda}$ includes three types of actions that give rise to the actions of $\pi_\Lambda$.
\begin{enumerate}
\item Actions for {\em inserting} a component (precondition, positive effect or negative effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models. These actions will form the prefix of the solution plan $\pi_\Lambda$. Among the \emph{inserting} actions, we find:
\begin{itemize}
\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$. A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $del_p$ nor $add_p$ exist in $\xi$.

@@

\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$. A positive effect is inserted in $\xi$ under the same conditions of a precondition insertion, and a negative effect is inserted in $\xi$ when neither $del_p$ nor $add_p$ appear in $\xi$ but $pre_p$ does.

@@

For instance, given $name(\xi)=${\tt{\small stack}} and $C_{pre-stack}=\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_holding\_v2), (pre\_stack\_on\_v1\_v2),(pre\_stack\_clear\_v1),(pre\_stack\_clear\_v1),}}$\ldots \}$, the insertion of each item $c \in C_{pre-stack}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$ as long as $c \notin pre(\xi)$,  $c \notin add(\xi)$ and $c \notin del(\xi)$. The same applies to effects with respect to sets $C_{add-stack}$ and $C_{del-stack}$ that would include all fluents starting with prefix  {\tt{\small add}} and {\tt{\small del}}, respectively.

@@

Note that executing an insert action, e.g.{\tt{\small (insert\_pre\_stack\_holding\_v1)}}, will add the corresponding model fluent {\tt{\small (pre\_stack\_holding\_v1)}} to the successor state. Hence, the execution of the insert actions of $\pi_\Lambda$ yield a state containing the valuation of the model fluents that shape every $\xi \in \mathcal{M}$. For example, executing the insert actions that shape the action model $name(\xi)=${\tt{\small putdown}} leads to a state containing the positive literals {\tt{\small (pre\_putdown\_holding\_v1),(eff\_putdown\_holding\_v1),\\ (eff\_putdown\_clear\_v1),
(eff\_putdown\_ontable\_v1),(eff\_putdown\_handempty)}}.

@@

    These actions will be part of the postfix of the plan $\pi_\Lambda$ and they determine the application of the learned action models according to the values of the model fluents in the current state configuration. Figure~\ref{fig:compilation} shows the PDDL encoding of {\tt{\small (apply\_stack)}} for applying the action model of the {\em stack} operator. Let's assume the action {\tt{\small (apply\_stack blockB blockA)}} is in $\pi_\Lambda$. Executing this action in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$. For example, if  $\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}}$\} \subset s$ then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$. Otherwise, a different set of precondition literals will be checked. The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.

Note that executing an apply action, e.g.{\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if $name(\xi)=${\tt{\small stack}} has been correctly programmed by the insert actions. Hence, while \textbf{insert actions} add the values of the \textbf{model fluents} that shape $\xi$, the \textbf{apply actions} add the values of the \textbf{fluents of $F$} that result from the execution of $\xi$.

@@

There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$. The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$.

@@

@@

The logical inference process our approach is based on has trouble learning preconditions that do not appear as negative effects since in this case no change is observed between the pre-state and post-state of an action. This is specially relevant for static predicates that never change and, hence, only appear as preconditions in the actions. In order to address this shortcoming and complete the list of learned preconditions, we apply a post-process based on the one proposed in~\cite{kuvcera2018louga}. The idea lies in going through every action and counting the number of cases where a literal is present before the action is executed and the number of cases where it is not present. If a literal is present in all the cases before the action, the literal is considered to be a precondition.

In order to obtain a complete trace, the proposal in ~\cite{kuvcera2018louga} applies the sequence of actions of the input trace and infers the preconditions from this \FO action sequence. In our case, since the sequence of actions of the input trace might not be fully observable, we produce the traces by applying the actions found in the validation part of the solution plan. For instance, in the example of the figure \ref{fig:plan-lplan}, the sequence of actions used to produce the complete trace would be {\tt{\small(unstack blockB blockA)}}, {\tt{\small(put-down blockB)}}, {\tt{\small(pick-up blockA)}}, and {\tt{\small(stack blockA blockB)}}.

@@

@@

\begin{proof}[Proof sketch]
\begin{small}
  By definition, $\Psi_{\xi}\subseteq \Psi_v$ fully captures the set of elements that can appear in an action model $\xi\in\mathcal{M}$. The compilation does not discard any possible set of action models $\mathcal{M}'$ definable within $\Psi_v$ that satisfies the observed state trajectory and action sequence of $\tau$. This means that for every $\mathcal{M}'$ that solves $\Lambda$, there exists a plan $\pi_\Lambda$ that can be built selecting the appropriate programming, apply and validate actions from the $P_{\Lambda}$ compilation.
\end{small}
\end{proof}

@@
